project-plan

### 2025-09-25 project-plan.qmd
Adding section about WHO:s TQIP guidelines into intro

@@

Despite the presence of systems and guidelines in place for these patients, trauma care remains vulnerable to errors,
deviations from established practice, and treatment delays, all of which may negatively influence outcomes. In
response, the American College of Surgeons developed the Trauma Quality Improvement Program (TQIP), a risk-adjusted
benchmarking system designed to assess the quality of trauma care. Through audit filters and peer review, it
identifies opportunities for improvement (OFI) and ultimately reduces trauma-related mortality (10,11). Previous
studies and reviews have discussed preventable trauma deaths where recurring patterns occur, namely, delays in
imaging and care, missed injuries, deviations in care regarding airway-related care, haemorrhage control, and fluid
management (12–16). Importantly, some of these findings were observed in high-performing level I trauma centres,
underscoring that even mature systems remain vulnerable to error. Furthermore, injuries to the central nervous system
account for a substantial proportion of trauma-related deaths, as demonstrated in analyses from large registries such
as the Arkansas Trauma Registry (16). Yet, while previous research has described OFI in trauma resuscitation and in
general ICU trauma cohorts, relatively little is known about the prevalence, nature, and determinants of OFI
specifically among patients with TBI. <!-- Add a part after ACS TQIP using WHO as a reference to TQIP-->

@@

@@

Furthermore, the categorical variables include…
<!-- Choose variables, not sure which too chose-->

@@

For continuous variables, medians and interquartile ranges (25th-75th  percentiles) are used. Categorical variables
are presented using the number of patients and their corresponding percentages of the cohort. <!-- If multivariable regression and anlysis will be used, how will we handle missing data? -->

Logistic regression was used to examine the association between patient and process-related factors and OFI. Results
are reported as Odds Ratio (OR) with 95% confidence interval (CI), and a p-value of <0.05 was used as the threshold
for statistical significance. To ensure objectivity, all statistical analyses were initially performed on synthetic
data before being applied to the combined trauma registry and trauma care quality database. Statistical analyses were
conducted in R version x.

<!-- unsure how the statistical model will work, previous papers mention multivariable logistic regression and one Uni/bivariable regression. How does this work? -->

>>>>

<!--NEED TO ADD SOURCES IN TEXT-->

@@

Despite the presence of systems and guidelines in place for these patients, trauma care remains vulnerable to errors,
deviations from established practice, and treatment delays, all of which may negatively influence outcomes. In
response, the American College of Surgeons developed the Trauma Quality Improvement Program (TQIP), a risk-adjusted
benchmarking system designed to assess the quality of trauma care. Through audit filters and peer review, it
identifies opportunities for improvement (OFI) and ultimately reduces trauma-related mortality (10,11). The WHO, however, insist that the trauma quality improvement programmes must be a continuous, multidisciplinary process. This
entails audit filters, structured morbidity and mortality and preventable death case reviews in order to find
opportunities for improvement (OFI). They also discuss that the formulation of corrective strategies or action plans
in response to identified OFI:s is essential to trauma quality improvement programmes. Furthermore, the WHO
emphasises that TQIP must ensure that corrective measures addressing identified OFI result in tangible and verifiable
improvements in outcomes, thereby confirming that deficiencies are not only recognised but also effectively resolved. (Källa WHO) <!-- add WHO source -->

Previous studies and reviews have discussed preventable trauma deaths where recurring patterns occur, namely, delays in imaging and care, missed injuries, deviations in care regarding airway-related care, haemorrhage control, and fluid management (12–16). Importantly, some of these findings were observed in high-performing level I trauma centres, underscoring that even mature systems remain vulnerable to error. Furthermore, injuries to the central nervous system account for a substantial proportion of trauma-related deaths, as demonstrated in analyses from large registries such as the Arkansas Trauma Registry (16). Yet, while previous research has described OFI in trauma resuscitation and in general ICU trauma cohorts, relatively little is known about the prevalence, nature, and determinants of OFI specifically among patients with TBI. <!-- Add a part after ACS TQIP using WHO as a reference to TQIP-->

@@

Aim
=====
As such, this study aims to identify and describe opportunities for improvement in the care of adult trauma patients
with TBI, and to examine patient- and process-related factors associated with their occurrence.

@@

Furthermore, the categorical variables include…
<!-- Choose variables, not sure which ones to chose-->

@@

For continuous variables, medians and interquartile ranges (25th-75th  percentiles) are used. Categorical variables
are presented using the number of patients and their corresponding percentages of the cohort. <!-- If multivariable regression analysis will be used, how will we handle missing data from other variables in the same patient? -->

Logistic regression was used to examine the association between patient and process-related factors and OFI.
Bivariable logistic regression to assess unadjusted associations between patient and process factors and OFI was
initially performed. Variables of clinical relevance and those significant in bivariable analyses were then entered
into a multivariable logistic regression model to estimate adjusted associations. Results are reported as Odds Ratio
(OR) with 95% confidence interval (CI), and a p-value of <0.05 was used as the threshold for statistical
significance. To ensure objectivity, all statistical analyses were initially performed on synthetic data before being
applied to the combined trauma registry and trauma care quality database. Statistical analyses were conducted in R
version x. <!-- I assume we will do both bivariable and multivariable regression? -->
